---
title: "preliminary_analyses"
author: "ML"
date: "08/06/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(DT)
library(ggimage)
library(ggcharts)
library(scales)

## set a plotting theme
theme_set(theme_classic(base_size = 14) + 
            theme(panel.background = element_blank(), 
                  plot.background = element_blank(), 
                  legend.background = element_blank(), 
                  text = element_text(colour = "white")))

#check oout other themes: https://r-charts.com/ggplot2/themes/
# library(ggthemr)
# ggthemr("solarized")
# ggthemr("freshe")
# ggthemr("pale")
 #call ggthemr_reset() #to reset to the default ggplot2 theme
```



## BR awards

### Load and process data

```{r BR load and clean data}
BRdata <- read_csv(file = here("data", "Copy of Ecoevo_awards_survey-Best_researcher (final) - Form Responses 1.csv"))
str(BRdata)
BRdata <- select(BRdata, -c(1:2)) #remove first two columns (non-informative)
#BR_orig_colnames <- names(BRdata) #store original column names

#rename selected data columns
BRdata <- BRdata %>% 
rename(Award_name = "Full name of the award", 
       Society_name = "Full name of the awarding society", 
       Disciplinary_focus = "Disciplinary focus",
       Geographical_range = "Geographical range", 
       EDI_policies = "Commitment to EDI in the society policies",
       EDI_structures = "Commitment to EDI in the society structures",
       Career_stage = "Target career stage of eligible applicants, as stated in the award information",
       Flexible_eligibility = "Flexibility of the eligibility criteria", 
       Inclusivity_statement = "Inclusivity statement", 
       Assessors_transparency = "Assessors transparency", 
       Process_transparency = "Process transparency",
       Feedback_availability = "Feedback availability",
       Criteria_transparency = "Criteria transparency",
       Relative_opportunity = "Assessment relatively to opportunity",  
       Diverse_contributions = "Valuing diverse contributions",    
       Open_science = "Valuing Open Science",  
       Self_nomination = "Self-nomination allowed",
       Letter_required = "Letter required", 
       Awardee_years = "Awardee list number of years",  
       N_female_2011_2020 = "Number of female awardees 2011-2020", 
       N_male_2011_2020 = "Number of male awardees 2011-2020",  
       N_missing_2011_2020 = "Number of awardees with unassignable gender 2011-2020",
       N_female_2001_2010 = "Number of female awardees 2001-2010", 
       N_male_2001_2010 = "Number of male awardees 2001-2010", 
       N_missing_2001_2010 = "Number of awardees with unassignable gender 2001-2010",
       N_female_1991_2000 = "Number of female awardees 1991-2000", 
       N_male_1991_2000 = "Number of male awardees 1991-2000",  
       N_missing_1991_2000 = "Number of awardees with unassignable gender 1991-2000", 
       N_female_1981_1990 = "Number of female awardees  1981-1990",
       N_male_1981_1990 = "Number of male awardees  1981-1990",
       N_missing_1981_1990 = "Number of awardees with unassignable gender 1981-1990", 
       N_more_data = "More awardee data"
       )
#names(BRdata)

#create a new column Award by combining society and award names:
BRdata %>% 
  mutate(Award = case_when(Award_name == "Open Science in Practice" ~ "SORTEE Open Science in Practice",
                           Award_name == "John Maynard Smith Prize" ~ "ESEB Maynard Smith Prize",
                           Award_name == "President's Award" ~ "ESEB President's Award",
                           Award_name == "T. Dobzhansky Prize" ~ "SSE Dobzhansky Prize",
                           Award_name == "ASN Distinguished Naturalist Award (previously E. O. Wilson Naturalist Award)" ~ "ASN Distinguished Naturalist Award",
                           Award_name == "ASN Early Career Investigator Awardn (Previously Jasper Loftus-Hills Young Investigator Award, the Young Investigator Prize, and the Jasper Loftus-Hills Young Investigator Prize)" ~ "ASN Early Career Investigator Award",
                           Award_name == "GfÖ-Prize" ~ "ESGAS GfÖ-Prize",
                           Award_name == "SMBE Early-Career Excellence Award" ~ "SMBE Early-Career Excellence Award",
                           Award_name == "SMBE Mid-Career Excellence Award" ~ "SMBE Mid-Career Excellence Award",
                           Award_name == "The Bicentenary Medal" ~ "LSL The Bicentenary Medal",
                           Award_name == "International Recognition of Professional Excellence" ~ "IEI Professional Excellence",
                           Award_name == "Robert H. MacArthur Award" ~ "ESA MacArthur Award",
                           Award_name == "Founders’ Prize" ~ "BES Founders’ Prize"
                           )
         ) -> BRdata

#Create reverse values on Letter_not_required - so "yes" is better
BRdata %>% 
  mutate(Letter_not_required = case_when(Letter_required == "yes" ~ "no",
                           Letter_required == "no" ~ "yes",
                           Letter_required == "uncelar" ~ "unclear",
                           )) -> BRdata


#Merge and shorten Society and Award names for plotting:
BRdata %>% 
  mutate(Award = case_when(Award_name == "Open Science in Practice" ~ "SORTEE Open Science in Practice",
                           Award_name == "John Maynard Smith Prize" ~ "ESEB Maynard Smith Prize",
                           Award_name == "President's Award" ~ "ESEB President's Award",
                           Award_name == "T. Dobzhansky Prize" ~ "SSE Dobzhansky Prize",
                           Award_name == "ASN Distinguished Naturalist Award (previously E. O. Wilson Naturalist Award)" ~ "ASN Distinguished Naturalist Award",
                           Award_name == "ASN Early Career Investigator Awardn (Previously Jasper Loftus-Hills Young Investigator Award, the Young Investigator Prize, and the Jasper Loftus-Hills Young Investigator Prize)" ~ "ASN Early Career Investigator Award",
                           Award_name == "GfÖ-Prize" ~ "ESGAS GfÖ-Prize",
                           Award_name == "SMBE Early-Career Excellence Award" ~ "SMBE Early-Career Excellence Award",
                           Award_name == "SMBE Mid-Career Excellence Award" ~ "SMBE Mid-Career Excellence Award",
                           Award_name == "The Bicentenary Medal" ~ "LSL The Bicentenary Medal",
                           Award_name == "International Recognition of Professional Excellence" ~ "IEI Professional Excellence",
                           Award_name == "Robert H. MacArthur Award" ~ "ESA MacArthur Award",
                           Award_name == "Founders’ Prize" ~ "BES Founders’ Prize"
                           )
         ) -> BRdata

#BRdata$Award

```

### Tables

Society-related descriptive and categorical data - as a table

See: 
https://www.r-bloggers.com/2016/11/conditional-formatting-of-a-table-in-r/
https://www.littlemissdata.com/blog/prettytables
https://stackoverflow.com/questions/40871295/adding-logos-images-to-the-side-of-a-data-table

```{r BR table using formattable, eval=FALSE}
library(data.table)
library(formattable)

#prepare data - select columns with descriptive values on the awarding societies
BRdata %>% select(Award, Career_stage, Society_name, Disciplinary_focus, Geographical_range, EDI_policies, EDI_structures) -> BRdata1

#format the table
formattable(BRdata1, 
            align =c("l","c","c","c","c", "c", "c"), 
            list(`Award` = formatter("span", style = ~ style(color = "grey",font.weight = "bold"))
              ))

#NOTE: this table will not render in pdf? Also, difficult to do conditional formatting for factors
```


```{r BR table using DT, eval=TRUE}

#prepare data - select columns with descriptive values on the awarding societies
BRdata %>% select(Award, Society_name, Disciplinary_focus, Geographical_range, Career_stage, EDI_policies, EDI_structures) -> BRdata1

#format using functions from DT package:
datatable(BRdata1, options = list(pageLength = 15, autoWidth = TRUE)) %>% 
  formatStyle(columns = colnames(BRdata1), 
              backgroundColor = styleEqual(c("yes", "no", "unclear"), c("#009E73", "#D55E00", "#999999"))
             )

#based on: https://rstudio.github.io/DT/010-style.html

# unique(BRdata1$Society_name)
# table(BRdata1$Disciplinary_focus)
# table(BRdata1$Geographical_range)
# table(BRdata1$EDI_policies)
# table(BRdata1$EDI_structures)
# table(BRdata1$Career_stage)

write.csv(BRdata1, here("tables","BR_awards_table_v0.csv"))
```


### Simple plots

Plot a dot plot for columns with yes/no/unclear answers

```{r BR dot plot, echo=FALSE}
#prepare data - select columns with yes/no/unclear values
BRdata %>% select(Award, Flexible_eligibility, Inclusivity_statement, Assessors_transparency, Process_transparency, Feedback_availability, Criteria_transparency, Relative_opportunity, Diverse_contributions, Open_science, Self_nomination, Letter_not_required)  -> BRdata2

#reshape and set a custom order of factor levles for plotting (otherwise it will be reverse alphabetical order):
BRdata2 %>%
  gather(key = var_name, value = value, 2:ncol(BRdata2)) %>% 
  arrange(value) %>%
  mutate(var_name = factor(var_name, levels = rev(c("Flexible_eligibility", "Self_nomination", "Letter_not_required", "Inclusivity_statement", "Assessors_transparency", "Process_transparency", "Criteria_transparency",  "Diverse_contributions", "Open_science", "Relative_opportunity", "Feedback_availability")))) %>%
  ggplot(aes(y = var_name, x = Award, col = value)) +
  geom_point(size = 5) +
  scale_color_manual(values = c("#D55E00", "#999999", "#009E73")) +
  theme_minimal() +
  scale_x_discrete(position = "top") +
  theme(panel.grid.major = element_blank(), axis.text.x = element_text(angle = 45, hjust = 0),
        axis.title.x = element_blank(), axis.title.y = element_blank(), 
        panel.border = element_rect(colour = "black", fill = NA))

ggsave(here("plots","BR_dotplot_v0.png"), #plot = p1, 
       width = 15, height = 15, units = "cm", dpi = "retina")

ggsave(here("plots","BR_dotplot_v0.pdf"), device = "pdf", #plot = p1, 
       width = 15, height = 15, units = "cm", dpi = "retina")
```

For potentially using icons in the plot: 
#https://jtr13.github.io/cc21fall1/icons-and-symbols-in-r.html
#https://apps.timwhitlock.info/emoji/tables/unicode  
#p + ggimage::geom_emoji(aes(image = ifelse(abs(Petal.Length-fitted) > 0.5, '1f645', '1f600')), cex=0.04)


Plot the range of years for which awardee info is available

```{r BR plot award years, echo=FALSE}
BRdata %>%
  bar_chart(x = Award, y = Awardee_years) + 
  theme_classic() +
  ylim(0, 80) +
  ylab("Year range of awardee data") 

ggsave(here("plots", "BR_awards_years_range_v0.png"), width = 18, height = 10, units = "cm", dpi = "retina")
```

Plot awardee genders across award years

```{r BR plot awardees gender by years, echo=FALSE}
names(BRdata)
#Reshape data for plotting
BRdata %>% select(Award, 
                  N_female_2011_2020,
                  N_male_2011_2020, 
                  N_missing_2011_2020, 
                  N_female_2001_2010, 
                  N_male_2001_2010, 
                  N_missing_2001_2010, 
                  N_female_1991_2000, 
                  N_male_1991_2000, 
                  N_missing_1991_2000, 
                  N_female_1981_1990, 
                  N_male_1981_1990, 
                  N_missing_1981_1990)  -> BRdata3

## replace all NA with 0 (optional):
BRdata3[is.na(BRdata3)] <- 0

## reshape into long dataframe format:
BRdata3 %>% gather(key = var_name, value = value, 2:ncol(BRdata3)) -> BRdata3long

## create new variables coding gender and years:
BRdata3long %>% 
  mutate(Gender = case_when(startsWith(var_name, "N_female_") ~ "female",
                           startsWith(var_name, "N_male_") ~ "male",
                           startsWith(var_name, "N_missing") ~ "unassigned",
                           TRUE ~ "NA")) %>% 
  mutate(Years = case_when(endsWith(var_name, "2011_2020") ~ "2011-2020",
                           endsWith(var_name, "2001_2010") ~ "2001-2010",
                           endsWith(var_name, "1991_2000") ~ "1991-2000",
                           endsWith(var_name, "1981_1990") ~ "1981-1990",
                           TRUE ~ "NA")) -> BRdata3long

## set the levels in order we want:
BRdata3long$Gender <- factor(BRdata3long$Gender, levels = c("male", "female", "unassigned"))
#BRdata3long$Years <- factor(BRdata3long$Years, levels = c("2011-2020", "2001-2010", "1991-2000", "1981-1990"))
BRdata3long$Award <- factor(BRdata3long$Award, levels = rev(c("SORTEE Open Science in Practice","ESGAS GfÖ-Prize","LSL The Bicentenary Medal","SMBE Early-Career Excellence Award","SMBE Mid-Career Excellence Award","ESEB President's Award","BES Founders’ Prize","ASN Distinguished Naturalist Award","ESEB Maynard Smith Prize","IEI Professional Excellence","ASN Early Career Investigator Award","ESA MacArthur Award","SSE Dobzhansky Prize"))) #from oldest to newest

# ## plot for a single award - not stacked, with count values:
# BRdata3long %>% 
#   filter(Award == "ASN Early Career Investigator Award") %>% 
#   ggplot(aes(x = Years, y = value)) +
#   geom_bar(aes(fill = Gender), stat = "identity", position = position_dodge(0.9)) +
#   scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) +
#   theme_classic()

## plot for a all awards, as facets - not stacked, with count values:
BRdata3long %>% 
  filter(Award != "SORTEE Open Science in Practice") %>% 
  ggplot(aes(x = Years, y = value)) +
  geom_bar(aes(fill = Gender), stat = "identity", position = position_dodge(0.9)) +
  facet_wrap(~Award, scales = "fixed", nrow = 4, ncol = 3) +
  scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) +
  theme_classic() +
  theme(legend.position="top", legend.box = "horizontal", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Awardee name") +
  ylab("Count") 

ggsave(here("plots", "BR_awards_gender_years_counts.png"), width = 18, height = 20, units = "cm", dpi = "retina")


## plot for a all awards, as facets - not stacked, with percent values
BRdata3long %>% 
  filter(Award != "SORTEE Open Science in Practice") %>% 
  ggplot(aes(x = Years, y = value)) +
  geom_bar(aes(fill = Gender), stat = "identity", position = "fill") +
  facet_wrap(~Award, scales = "fixed", nrow = 4, ncol = 3) +
  scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) +
  geom_abline(slope = 0, intercept = 0.5, col = "red", lty = 2) + 
  theme_classic() +
  theme(legend.position="top", legend.box = "horizontal", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Awardee name") +
  ylab("Perccent") + 
  scale_y_continuous(labels = scales::percent)

ggsave(here("plots", "BR_awards_gender_years_v1.png"), width = 18, height = 20, units = "cm", dpi = "retina")


## plot for a all awards, as facets- not stacked, with percent values and count values
BRdata3long %>% 
  mutate(Award_year = paste(Award, Years)) %>%
  group_by(Award_year) %>% 
  mutate(Perc = value/sum(value)) %>% 
  filter(value != 0) %>% 
  ggplot(aes(Years, Perc, fill = Gender), stat = "identity", position = "fill") + 
  facet_wrap(~Award, scales = "fixed", nrow = 4, ncol = 3) +
  scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) + 
  geom_col(position = position_stack()) + 
  geom_text(aes(label = value), position = position_stack(vjust = 0.5), size = 4) +
  geom_abline(slope = 0, intercept = 0.5, col = "red", lty = 2) + 
  theme_classic() +
  theme(legend.position = "top", legend.box = "horizontal", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Awardee name") +
  ylab("Percent") + 
  scale_y_continuous(labels = scales::percent)

ggsave(here("plots", "BR_awards_gender_years_v2.png"), width = 18, height = 20, units = "cm", dpi = "retina")
```



## BP awards

### Load and process data

```{r BP load and clean data}
BPdata <- read_csv(file = here("data", "Copy of Ecoevo_awards_survey-Best_paper (final) - Form Responses 1.csv"))
str(BPdata)
names(BPdata)

BPdata <- select(BPdata, -c(1:2)) #remove first two columns (non-informative)
#BP_orig_colnames <- names(BPdata) #store original column names

#rename selected data columns
BPdata <- BPdata %>% 
rename(Award_name = "Full name of the award", 
       Journal_name = "Full name of the awarding journal", 
       Disciplinary_focus = "Journal disciplinary focus",
       Geographical_range = "Journal geographical range", 
       EDI_policies = "Journal commitment to EDI in the journal policies",
       Career_stage = "Target career stage of eligible applicants, as stated in the award information",
       Flexible_eligibility = "Flexibility of the eligibility criteria", 
       Inclusivity_statement = "Inclusivity statement", 
       Assessors_transparency = "Assessor transparency", 
       Process_transparency = "Process transparency",
       Feedback_availability = "Feedback availability",
       Criteria_transparency = "Criteria transparency",
       Open_science = "Valuing Open Science",  
       Self_nomination = "Self-nomination allowed",
       Letter_required = "Letter required", 
       Awardee_years = "Awardee list number of years",  
       N_female_2011_2020 = "Number of female awardees 2011-2020", 
       N_male_2011_2020 = "Number of male awardees 2011-2020",  
       N_missing_2011_2020 = "Number of awardees with unassignable gender 2011-2020",
       N_female_2001_2010 = "Number of female awardees 2001-2010", 
       N_male_2001_2010 = "Number of male awardees 2001-2010", 
       N_missing_2001_2010 = "Number of awardees with unassignable gender 2001-2010",
       N_female_1991_2000 = "Number of female awardees 1991-2000", 
       N_male_1991_2000 = "Number of male awardees 1991-2000",  
       N_missing_1991_2000 = "Number of awardees with unassignable gender 1991-2000", 
       N_female_1981_1990 = "Number of female awardees  1981-1990",
       N_male_1981_1990 = "Number of male awardees  1981-1990",
       N_missing_1981_1990 = "Number of awardees with unassignable gender 1981-1990", 
       N_more_data = "More awardee data"
       )
#names(BPdata)

#remove excluded award from Oecologia and create a new column Award by combining journal and award names: 
BPdata %>% filter(Journal_name != "Oecologia") %>% 
  mutate(Award = case_when(Journal_name == "Molecular Biology & Evolution; Society for Molecular Biology and Evolution" ~ "MolBiolEvol / SMBE Best Graduate Student Paper",
                           Journal_name == "Genome Biology & Evolution; Society for Molecular Biology and Evolution" ~ "GenomeBiolEvol / SMBE Best Graduate Student Paper",
                           Journal_name == "any journal; Ecological Society of America" ~ "any journal / ESA George Mercer Award",
                           Journal_name == "Journal of Evolutionary Biology; European Society for Evolutionary Biology" ~ "JEvolutionBiol / ESEB Stearns Graduate Student Prize",
                           Journal_name == "Evolution; Society for the Study of Evolution" ~ "Evolution / SSE Outstanding Dissertation Award",
                           Journal_name == "The American Naturalist; American Society of Naturalists" ~ "AmNat / ASN Student Paper Award",
                           Journal_name == "Methods in Ecology and Evolution; British Ecological Society" ~ "MethodsEcolEvol / BES Robert May Prize",
                           Journal_name == "Functional Ecology; British Ecological Society" ~ "FunctEcol / BES Haldane ECR Award",
                           Journal_name == "Ecology Letters; CNRS" ~ "EcolLett / CNRS ECR Award",
                           Journal_name == "Journal of Experimental Biology; The Company of Biologists" ~ "JExpBiol / CoB Outstanding Paper Prize"
                           )
         ) -> BPdata


#Create reverse values on Letter_not_required - so "yes" is better
BPdata %>% 
  mutate(Letter_not_required = case_when(Letter_required == "yes" ~ "no",
                           Letter_required == "no" ~ "yes",
                           Letter_required == "uncelar" ~ "unclear",
                           )) -> BPdata

#BPdata$Award

#set a custom order of Award levels for plotting 
BPdata <- BPdata %>% mutate(Award = factor(Award, levels = c("Evolution / SSE Outstanding Dissertation Award", 
                                                "AmNat / ASN Student Paper Award", 
                                                "EcolLett / CNRS ECR Award",
                                                "FunctEcol / BES Haldane ECR Award",
                                                "JEvolutionBiol / ESEB Stearns Graduate Student Prize",
                                                "JExpBiol / CoB Outstanding Paper Prize",
                                                "MethodsEcolEvol / BES Robert May Prize",
                                                "MolBiolEvol / SMBE Best Graduate Student Paper",
                                                "GenomeBiolEvol / SMBE Best Graduate Student Paper",
                                                "any journal / ESA George Mercer Award"
                                                )))
levels(BPdata$Award)
```

### Tables

Journal-related descriptive and categorical data - as a table

```{r BP table using , eval=TRUE}

#prepare data - select columns with descriptive values on the awarding societies
BPdata %>% select(Award, Journal_name, Disciplinary_focus, Geographical_range, Career_stage, EDI_policies) -> BPdata1

#format using functions from DT package:

datatable(BPdata1, options = list(pageLength = 15, autoWidth = TRUE)) %>% 
  formatStyle(columns = colnames(BPdata1), 
              backgroundColor = styleEqual(c("yes", "no", "unclear"), c("#009E73", "#D55E00", "#999999"))
             )

# unique(BPdata1$Journal_name)
# table(BPdata1$Disciplinary_focus)
# table(BPdata1$Geographical_range)
# table(BPdata1$EDI_policies)
# table(BPdata1$Career_stage)

write.csv(BPdata1, here("tables","BP_awards_table_v0.csv"))
```


### Simple plots

Plot a dot plot for columns with yes/no/unclear answers

```{r BP dot plot, echo=FALSE}
#prepare data - select columns with yes/no/unclear values
BPdata %>% select(Award, Flexible_eligibility, Inclusivity_statement, Assessors_transparency, Process_transparency, Feedback_availability, Criteria_transparency, Open_science, Self_nomination, Letter_not_required)  -> BPdata2

#reshape and set a custom order of factor levels for plotting (otherwise it will be reverse alphabetical order):
BPdata2 %>%
  gather(key = var_name, value = value, 2:ncol(BPdata2)) %>% 
  arrange(value) %>%
  mutate(var_name = factor(var_name, levels = rev(c("Flexible_eligibility", "Self_nomination", "Letter_not_required", "Inclusivity_statement", "Assessors_transparency", "Process_transparency", "Criteria_transparency", "Open_science", "Feedback_availability")))) %>%
  ggplot(aes(y = var_name, x = Award, col = value)) +
  geom_point(size = 5) +
  scale_color_manual(values = c("#D55E00", "#999999", "#009E73")) +
  theme_minimal() +
  scale_x_discrete(position = "top") +
  theme(panel.grid.major = element_blank(), axis.text.x = element_text(angle = 45, hjust = 0),
        axis.title.x = element_blank(), axis.title.y = element_blank(), 
        panel.border = element_rect(colour = "black", fill = NA),
        plot.margin = margin(t = 25,  # Top margin
                             r = 50,  # Right margin
                             b = 10,  # Bottom margin
                             l = 15)) # Left margin)

ggsave(here("plots","BP_dotplot_v0.png"), #plot = p1, 
       width = 15, height = 15, units = "cm", dpi = "retina")

ggsave(here("plots","BP_dotplot_v0.pdf"), device = "pdf", #plot = p1, 
       width = 15, height = 15, units = "cm", dpi = "retina")
```

Plot the range of years for which awardee info is available

```{r BP plot award years, echo=FALSE}
BPdata %>%
  bar_chart(x = Award, y = Awardee_years) + 
  theme_classic() +
  ylim(0, 80) +
  ylab("Year range of awardee data") 

ggsave(here("plots", "BP_awards_years_range_v0.png"), width = 18, height = 10, units = "cm", dpi = "retina")
```

Plot awardee genders across award years

```{r BP plot awardees gender by years, echo=FALSE}
#Reshape data for plotting
BPdata %>% select(Award, 
                  N_female_2011_2020,
                  N_male_2011_2020, 
                  N_missing_2011_2020, 
                  N_female_2001_2010, 
                  N_male_2001_2010, 
                  N_missing_2001_2010, 
                  N_female_1991_2000, 
                  N_male_1991_2000, 
                  N_missing_1991_2000, 
                  N_female_1981_1990, 
                  N_male_1981_1990, 
                  N_missing_1981_1990)  -> BPdata3

## replace all NA with 0 (optional):
BPdata3[is.na(BPdata3)] <- 0

## reshape into long dataframe format:
BPdata3 %>% gather(key = var_name, value = value, 2:ncol(BPdata3)) -> BPdata3long

## create new variables coding gender and years:
BPdata3long %>% 
  mutate(Gender = case_when(startsWith(var_name, "N_female_") ~ "female",
                           startsWith(var_name, "N_male_") ~ "male",
                           startsWith(var_name, "N_missing") ~ "unassigned",
                           TRUE ~ "NA")) %>% 
  mutate(Years = case_when(endsWith(var_name, "2011_2020") ~ "2011-2020",
                           endsWith(var_name, "2001_2010") ~ "2001-2010",
                           endsWith(var_name, "1991_2000") ~ "1991-2000",
                           endsWith(var_name, "1981_1990") ~ "1981-1990",
                           TRUE ~ "NA")) -> BPdata3long

## set the levels in order we want:
BPdata3long$Gender <- factor(BPdata3long$Gender, levels = c("male", "female", "unassigned"))
#BPdata3long$Years <- factor(BPdata3long$Years, levels = c("2011-2020", "2001-2010", "1991-2000", "1981-1990"))
BPdata3long$Award <- factor(BPdata3long$Award, levels = c("any journal / ESA George Mercer Award",
                                                          "Evolution / SSE Outstanding Dissertation Award",
                                                          "FunctEcol / BES Haldane ECR Award",
                                                          "AmNat / ASN Student Paper Award",
                                                          "JExpBiol / CoB Outstanding Paper Prize",
                                                          "MethodsEcolEvol / BES Robert May Prize",
                                                          "JEvolutionBiol / ESEB Stearns Graduate Student Prize",
                                                          "GenomeBiolEvol / SMBE Best Graduate Student Paper",
                                                          "MolBiolEvol / SMBE Best Graduate Student Paper",
                                                          "EcolLett / CNRS ECR Award")) #from longest to shortest years data

# ## plot for a single award - not stacked, with count values:
# BPdata3long %>% 
#   filter(Award == "ASN Early Career Investigator Award") %>% 
#   ggplot(aes(x = Years, y = value)) +
#   geom_bar(aes(fill = Gender), stat = "identity", position = position_dodge(0.9)) +
#   scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) +
#   theme_classic()

## plot for a all awards, as facets - not stacked, with count values:
BPdata3long %>% 
  ggplot(aes(x = Years, y = value)) +
  geom_bar(aes(fill = Gender), stat = "identity", position = position_dodge(0.9)) +
  facet_wrap(~Award, scales = "fixed", nrow = 5, ncol = 2) +
  scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) +
  theme_classic() +
  theme(legend.position="top", legend.box = "horizontal", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8), minor_breaks = NULL) +
  labs(fill = "Awardee name") +
  ylab("Count") 

ggsave(here("plots", "BP_awards_gender_years_counts.png"), width = 18, height = 20, units = "cm", dpi = "retina")


## plot for a all awards, as facets- not stacked, with percent values
BPdata3long %>% 
  ggplot(aes(x = Years, y = value)) +
  geom_bar(aes(fill = Gender), stat = "identity", position = "fill") +
  facet_wrap(~Award, scales = "fixed", nrow = 5, ncol = 2) +
  scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) +
  geom_abline(slope = 0, intercept = 0.5, col = "red",lty = 2) + 
  theme_classic() +
  theme(legend.position = "top", legend.box = "horizontal", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Awardee name") +
  ylab("Percent") + 
  scale_y_continuous(labels = scales::percent)


ggsave(here("plots", "BP_awards_gender_years_v1.png"), width = 18, height = 20, units = "cm", dpi = "retina")


## plot for a all awards, as facets- not stacked, with percent values and count values
BPdata3long %>% 
  mutate(Award_year = paste(Award, Years)) %>%
  group_by(Award_year) %>% 
  mutate(Perc = value/sum(value)) %>% 
  filter(value != 0) %>% 
  ggplot(aes(Years, Perc, fill = Gender), stat = "identity", position = "fill") + 
  facet_wrap(~Award, scales = "fixed", nrow = 5, ncol = 2) +
  scale_fill_manual(values = c("#56B4E9", "#E69F00", "#999999")) + 
  geom_col(position = position_stack()) + 
  geom_text(aes(label = value), position = position_stack(vjust = 0.5), size = 4) +
  geom_abline(slope = 0, intercept = 0.5, col = "red", lty = 2) + 
  theme_classic() +
  theme(legend.position = "top", legend.box = "horizontal", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Awardee name") +
  ylab("Percent") + 
  scale_y_continuous(labels = scales::percent)

ggsave(here("plots", "BP_awards_gender_years_v2.png"), width = 18, height = 20, units = "cm", dpi = "retina")

```
